---
title: "p8105_hw6_sef2183"
author: "Sarah Forrest"
date: "2022-11-17"
output: github_document
---

```{r, include = FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(purrr)
library(modelr)
library(mgcv)

knitr::opts_chunk$set(collapse = TRUE, message = FALSE, dpi = 300, fig.width = 7)

set.seed(1)
```

# Problem 1

Download the 2017 Central Park weather data using code adapted from the course website:
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

The boostrap is helpful when you’d like to perform inference for a parameter / value / summary that doesn’t have an easy-to-write-down distribution in the usual repeated sampling framework. We’ll focus on a simple linear regression with tmax as the response and tmin as the predictor, and are interested in the distribution of two quantities estimated from these data:

r̂ 2
log(β̂ 0∗β̂ 1)

Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities. Plot the distribution of your estimates, and describe these in words. Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for r̂ 2 and log(β̂ 0∗β̂ 1). Note: broom::glance() is helpful for extracting r̂ 2 from a fitted regression, and broom::tidy() (with some additional wrangling) should help in computing log(β̂ 0∗β̂ 1).

# Problem 2

The code below reads in the Washington Post's homicide data on homicides in 50 large U.S. cities. Then, a `city_state` variable (e.g. “Baltimore, MD”) and a binary variable `homicide_resolved` indicating whether the homicide is solved ("Yes" or "No") are created. Dallas, TX, Phoenix, AZ, and Kansas City, MO are omitted from the dataframe as these cities don’t report victim race. Tulsa, AL is also omitted from the dataframe due to a data entry mistake. Finally, the dataframe is limited to homicides for whom `victim_race` is white or black and the `victim_age` variable is converted to numeric:

```{r}
homicide_df = 
  read.csv("https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv") %>%
  mutate(
    city_state = paste(city, state, sep = ", "),
    homicide_resolved = as.numeric(disposition == "Closed by arrest")) %>%
  filter(city_state != "Dallas, TX", 
         city_state != "Phoenix, AZ",
         city_state != "Kansas City, MO",
         city_state != "Tulsa, AL",
         victim_race == "White" | victim_race == "Black") %>%
  mutate(victim_age = as.numeric(victim_age))
```

## Logistic Regression Output for Baltimore, MD

The code below uses the `glm()` function to fit a logistic regression with resolved vs. unresolved as the outcome and victim age, sex and race as predictors for the city of Baltimore, MD. The output of this regression saved as an R object called `fit_logistic`. 

```{r}
fit_logistic =
  homicide_df %>%
  filter(city_state == "Baltimore, MD") %>%
  glm(homicide_resolved ~ victim_age + victim_sex + victim_race, data = ., family = binomial())
```

The code below obtains the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed:

```{r}
fit_logistic %>%
  broom::tidy() %>%
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - (1.96 * std.error)),
    CI_upper = exp(estimate + (1.96 * std.error))) %>%
  select(term, OR, CI_lower, CI_upper) %>% 
  filter(term == "victim_sexMale") %>% 
  knitr::kable(digits = 2)
```

## Logistic Regression Output for All Cities

The code below creates a nested dataframe `homicide_nest_df` that contains data on whether each homicide is resolved, as well as victim demographic information for each city in the dataset. Then, a function called “glm_OR_CI” is created that takes in a dataframe and applies the `glm()` function to run a logistic regression model with resolved vs unresolved as the outcome and victim age, sex and race as predictors and calculate the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed. This function is called and mapped to a new `glm_output` variable which is then unnested and assigned to a new `homicide_glm_df` dataframe. The `homicide_glm_df` dataframe contains 5 variables: `city_state`, a `data` variable containing a tibble with the victim and homicide data used in the logistic regression model, `OR`, and the lower and upper confidence interval values `CI_lower` and `CI_upper`:

```{r}
homicide_nest_df = 
  homicide_df %>% 
  select(city_state, homicide_resolved, victim_race:victim_sex) %>%
  group_by(city_state) %>%
  nest(data = homicide_resolved:victim_sex)
  

glm_OR_CI = function(df) {
  
  fit_logistic = 
    glm(homicide_resolved ~ victim_age + victim_sex + victim_race, data = df, family = binomial())
  
  fit_logistic %>%
  broom::tidy() %>%
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - (1.96 * std.error)),
    CI_upper = exp(estimate + (1.96 * std.error))) %>%
  filter(term == "victim_sexMale") %>% 
  select(OR, CI_lower, CI_upper) 
  
}


homicide_nest_df %>% 
  mutate(glm_output = map(data, glm_OR_CI)) %>% 
  unnest(glm_output) -> homicide_glm_df

homicide_glm_df
```

## Plot 

The code below creates a plot that shows the estimated ORs and CIs for each city. Cities are organized according to estimated OR:

```{r}
homicide_glm_df %>%
  ggplot(aes(x = reorder(city_state, OR), y = OR)) +
  geom_point() +
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper), width = 0.2) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(
    title = "Adjusted OR for Solving Homicides Comparing Male to Female Victims",
    x = "City",
    y = "Adjusted OR Estimate")
```

The city with the smallest adjusted OR estimate for solving homicides comparing male victims to female victims is New York, NY. The confidence interval range for this estimate is fairly small, ranging from about 0.14 to 0.50. The city with the largest adjusted OR estimate is Albuquerque, NM. The confidence interval range for Albuquerque's OR estimate is much larger than New York, ranging from about 0.83 to 3.76. A general trend that can observed from the plot is that the confidence interval is wider as the estimated adjusted OR estimate increases.

# Problem 3

Read in the "birthweight" dataset consisting of several variables of data on roughly 4000 children:

```{r}
birthweight_df = read.csv("data/birthweight.csv")
```

## Create Model

The code below produces a smooth regression model for predictors of birthweight. The model includes birthweight as the outcome, and predictors of birthweight identified by Johansson et al., including: mother’s weight gain during pregnancy, average number of cigarettes smoked per day during pregnancy, mother’s pre-pregnancy BMI, number of live births prior to this pregnancy (parity), mother’s age at delivery, and baby’s sex:

Reference: 
Johansson K, Linné Y, Rössner S, Neovius M. Maternal predictors of birthweight: The importance of weight gain during pregnancy. Obes Res Clin Pract. 2007 Dec;1(4):223-90. doi: 10.1016/j.orcp.2007.09.001. PMID: 24351583.

```{r}
smooth_mod = mgcv::gam(bwt ~ s(wtgain + smoken + ppbmi + parity + momage + babysex), data = birthweight_df)
```

The code below produces a plot of model residuals against fitted values using `add_predictions()` and `add_residuals()`:

```{r}
birthweight_df %>% 
  add_predictions(smooth_mod) %>% 
  add_residuals(smooth_mod) %>% 
  ggplot(aes(x = pred, y = resid)) + 
  geom_point() 
```

## Compare Models

Comparison model 1: the code below produces a regression model for birthweight using length at birth and gestational age as predictors (main effects only):

```{r}
comp_mod1 = mgcv::gam(bwt ~ s(blength + sgaweeks), data = birthweight_df)
```

Comparison model 2: the code below produces a regression model for birthweight using head circumference, length, sex, and all interactions (including the three-way interaction) between these:

```{r}
comp_mod2 = mgcv::gam(bwt ~ s(bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead*blength*babysex), data = birthweight_df)
```

The code below compares models in terms of the cross-validated prediction error using `crossv_mc()` and `map()`. A boxplot for each model is created to compare the root-mean-square error (RMSE) values for each model:

```{r, warning = FALSE}
cv_df = 
  crossv_mc(birthweight_df, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)) %>% 
  mutate(
    smooth_mod = map(.x = train, ~mgcv::gam(bwt ~ s(wtgain + smoken + ppbmi + parity + momage + babysex), data = as_tibble(.x))),
    comp_mod1 = map(.x = train, ~mgcv::gam(bwt ~ s(blength + sgaweeks), data = as_tibble(.x))),
    comp_mod2 = map(.x = train, ~mgcv::gam(bwt ~ s(bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead*blength*babysex), data = as_tibble(.x)))) %>% 
  mutate(
    rmse_smooth_mod = map2_dbl(.x = smooth_mod, .y = test, ~rmse(model = .x, data = .y)),
    rmse_comp_mod1 = map2_dbl(.x = comp_mod1, .y = test, ~rmse(model = .x, data = .y)),
    rmse_comp_mod2 = map2_dbl(.x = comp_mod2, .y = test, ~rmse(model = .x, data = .y)))

cv_df

cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) %>% 
  ggplot(aes(x = model, y = rmse)) +
  geom_boxplot()
```

As seen in the boxplots above, the model with the lowest RMSE is comparison model 1 (the regression model for birthweight using length at birth and gestational age as predictors). Since te RMSE for this model is the lowest, this is the best fitting regression model for birthweight out of the 3 models that were compared. 